\chapter{Summer School on Effective HPC for Climate and Weather -- 2021}
\label{ch:ss2021}

\begin{comment}

\title{}

\section{}

\subsection{Registration}
\label{sec:registration}

The applicants also had to choose between two main packages representing the fee structure of the event:

\textbf{The Full Package Registration: £800}

\begin{itemize}

\item Registration in the event Mailing List
\item Attendance to all sessions and social events
\item Accommodation on university premises (Check Section Accommodation)
\item All-inclusive meals (breakfast, lunch, and three-course self-service dinner with tea and coffee)
\item Transport between London Heathrow Airport and the venue (by bus)

\end{itemize}

\textbf{The Minimal Registration: £350}

\begin{itemize}

\item Registration in the event Mailing List
\item Attendance to all sessions and social events
\item Lunch

\end{itemize}

\subsection{Funding of Attendees}
\label{sec:funding}

A committee will conduct the selection procedure for subsidising attendees, which will be funded according to their financial situation, the support provided by their home institution, and other possible fundings. We appreciate partial financing.

During the selection procedure, the applicants that request subsidy will be classified into one of the following categories:

\begin{description}

\item[Travel Subsidy] The Full Registration fee for the summer school will be covered. In addition, attendees will receive the reimbursement of their travel costs to Reading after the event. The receipts must be provided for the payment to be made in full. Note that we have only a few travel subsidy grants, so please attempt to support your funding.

\item[Attendance Subsidy] The Full Registration fee for the summer school will be included.
Registration Subsidy: The Minimal Registration fee for the summer school will be included.

\item[No Subsidy] Attendees will have to pay for the total cost of the summer school.
In all the cases in which a subsidy is granted, a fee prepayment (also described as a deposit) with a value of £100 will be charged to secure the participant's place on the event 4). Please, note that none of the subsidy categories covers personal expenses.

\end{description}

In addition to their (application process), applicants who want to be considered for subsidies will have to submit:

A one-page document describing:

\begin{itemize}

\item the justification for the subsidisation,
\item the requested level of subsidy, and
\item a tentative cost estimate (for travel subsidy).

\end{itemize}

Applicants will receive an answer from the committee four months before the event.

The form provided means to attach the letters (Session \ref{sec:letters}), indicate the type of registration (Session \ref{sec:registration}), and  whether a subsidy was required (Session \ref{sec:funding}).

For statistical purposes, we collected data on the applicants background in the fields explored in the summer school.

\end{comment}